---
title: "coat_to_camtrapdp"
author: "Julia Wiel"
date: "`r Sys.Date()`"
output: "html_document"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(here)
library(dplyr)
library(tidyr)
library(frictionless)
library(camtrapdp)
library(lubridate)
library(readr)
library(fs)
```

# Import metadata and classifications from borgefjell

```{r}
# Enter variables
year <- 2024

# Import files
setwd("/run/user/210015540/gvfs/sftp:host=login.nird.sigma2.no,user=jwiel/nird/projects/NS11082K/national_monitoring/")
metadata_raw <- readr::read_delim(here::here("gbif", "raw", "image_metadata", "national_monitoring_rodents_cameratraps_image_metadata_borgefjell_2024.txt"),
 delim = ";")

classif_raw <- readr::read_delim(here::here("gbif", "raw", "image_classification_small_mammal_model_2025", "national_monitoring_rodents_cameratraps_image_classification_borgefjell_2024.txt"),
 delim = ";")

coord_raw <- readr::read_delim(here::here("gbif", "raw", "Koordinates_SN_FI_RE_JO_KO_v2.csv"),
 delim = ";")

```

# Create resources

## Inspect raw data

```{r}
(unique_values_per_column <-
  metadata_raw %>%
  dplyr::summarise(dplyr::across(dplyr::everything(), ~ dplyr::n_distinct(.))) %>%
  tidyr::pivot_longer(dplyr::everything(), names_to = "column", values_to = "n_unique"))

(unique_values_per_column <-
  classif_raw %>%
  dplyr::summarise(dplyr::across(dplyr::everything(), ~ dplyr::n_distinct(.))) %>%
  tidyr::pivot_longer(dplyr::everything(), names_to = "column", values_to = "n_unique"))

(unique_values_per_column <-
  coord_raw %>%
  dplyr::summarise(dplyr::across(dplyr::everything(), ~ dplyr::n_distinct(.))) %>%
  tidyr::pivot_longer(dplyr::everything(), names_to = "column", values_to = "n_unique"))
```

## Format information from metadata

```{r}
metadata <- metadata_raw

# Format captureMethod
metadata$captureMethod <- metadata %>%
 dplyr::mutate(captureMethod = dplyr::case_when(
  v_sequence == 0 ~ "timeLapse",
  v_sequence > 0  ~ "activityDetection",
  TRUE            ~ NA_character_
))

# Format timestamp to UTC
metadata$timestamp <- format(lubridate::ymd_hms(paste(metadata$t_date,metadata$t_time," "), tz = "CEST"), 
format = "%Y-%m-%dT%H:%M:%OSZ")

 # Format deployment dates
 metadata <- metadata %>%
 dplyr::mutate(deploymentStart = min(timestamp),
 deploymentEnd = max(timestamp))

 ## Add coordinates to metadata
 # Transform the key column
coord <- coord_raw
coord <- coord %>%
  dplyr::mutate(sn_site = tolower(name)) %>% 
  mutate(sn_site = gsub("-", "_", sn_site))

# Merge columns from Koordinates into metadata
metadata <- metadata %>% 
  dplyr::left_join(coord, by = "sn_site")

# Create media filePath
metadata <- metadata %>%
  dplyr::mutate(filePath = fs::path("images_renamed", sn_locality, sn_site, year,v_image_name))

```

## Format information from classifification

```{r}
classif <- classif_raw

# Add observationType
animal <- c("aves", "cricetidae", "lem_lem", "mus_erm", "mus_niv", "sor_sp", "unidentified_animal")
classif <- classif %>%
  dplyr::mutate(observationType = dplyr::case_when(
    v_class_id %in% animal      ~ "animal",
    v_class_id == "bad_quality" ~ "unknown",
    v_class_id == "empty"       ~ "blank",
    TRUE                        ~ "unclassified"
  ))

# Add scientificName
classif <- classif %>%
  dplyr::mutate(scientificName = dplyr::case_when(
    v_class_id == "aves"     ~ "Aves",
    v_class_id == "cricetidae" ~ "Cricetidae",
    v_class_id == "lem_lem"       ~ "Lemmus lemmus",
    v_class_id == "mus_erm"       ~ "Mustela erminea",
    v_class_id == "mus_niv"      ~ "Mustela nivalis",
    v_class_id == "sor_sp"      ~ "Sorex spp.",
    TRUE                        ~ NA_character_
  ))
```

## Create deployments resource_start changes

```{r}
deployments <-
  metadata %>% 
  dplyr::mutate(
    deploymentID = metadata$sn_site, # Required
    locationID = metadata$sc_section,
    locationName = metadata$sn_locality,
    latitude = metadata$lat, # Required
    longitude = metadata$lon, # Required
    coordinateUncertainty = NA_integer_,
    deploymentStart = metadata$deploymentStart, # Required
    deploymentEnd = metadata$deploymentEnd, # Required
    setupBy = NA_character_,
    cameraID = NA_character_,
    cameraModel = NA_character_,
    cameraPlatform = NA_character_,
    cameraDelay = 60, # Eivind confirms?
    cameraHeight = NA_real_,
    cameraDepth = NA_real_,
    cameraTilt = -90,
    cameraHeading = NA_integer_,
    detectionDistance = NA_real_,
    timestampIssues = NA,
    baitUse = FALSE,
    featureType = NA_character_,
    habitat = metadata$sc_type_of_sites_ecological, # Or "habitat" from coord_raw?
    deploymentGroups = NA_character_,
    deploymentTags = NA_character_,
    deploymentComments = NA_character_,
    .keep = "none") %>% 
  dplyr::arrange(deploymentID) %>% 
  dplyr::select(
    deploymentID, locationID, locationName, latitude, longitude, 
    coordinateUncertainty, deploymentStart, deploymentEnd, setupBy, 
    cameraID, cameraModel, cameraPlatform, cameraDelay, cameraHeight, 
    cameraDepth, cameraTilt, cameraHeading, detectionDistance,
    timestampIssues, baitUse, featureType, habitat, deploymentGroups,
    deploymentTags, deploymentComments
  )
```

## Create media resource

```{r}
media <-
  metadata %>% 
  dplyr::mutate(
    mediaID = metadata$v_image_name, # Required
    deploymentID = metadata$sn_site, # Required
    captureMethod = metadata$captureMethod,
    timestamp = metadata$timestamp, # Required 
    filePath = NA_character_,# Required
    filePublic = FALSE, # Required
    fileName = metadata$v_image_name_original,
    fileMediatype = "image/jpeg", # Required
    exifmetadata = NA, # maybe reconix https://camtrap-dp.tdwg.org/metadata/#media.exifmetadata
    favorite = NA,
    mediaComments = metadata$v_comment,
    .keep = "none"
  ) %>% 
  dplyr::arrange(mediaID) %>%
  dplyr::distinct() %>% 
  dplyr::select(
    mediaID, deploymentID, captureMethod, timestamp, filePath, 
    filePublic, fileName, fileMediatype, exifmetadata, favorite, mediaComments
  )
```

## Create observations resource

```{r}
observations <-
  classif %>% 
  dplyr::mutate(
    observationID = NA_character_, # Required
    deploymentID = classif$sn_site, # Required
    mediaID = classif$v_image_name,
    eventID = NA_character_, # maybe metadata$v_event
    eventStart = NA_Date_, # Required Earliest event time stamp for a same v_event
    eventEnd = NA_Date_, # Required Latest event time stamp for a same v_event
    observationLevel = "media", # Required
    observationType = classif$observationType, # Required
    cameraSetupType = NA_character_,
    scientificName = classif$scientificName,
    count = NA_integer_,
    lifeStage = NA_character_,
    sex = NA_character_,
    behavior = NA_character_,
    individualID = NA_character_,
    individualPositionRadius = NA_real_,
    individualPositionAngle = NA_real_,
    individualSpeed = NA_real_,
    bboxX = NA_real_,
    bboxY = NA_real_,
    bboxWidth = NA_real_,
    bboxHeight = NA_real_,
    classifificationMethod = "machine", # or split manual and automatic
    classififiedBy = "Small_mammal_classification_model_version_2025",
    classifificationTimestamp = NA_Date_,
    classifificationProbability = classif$v_confidence_automatic,
    observationTags = NA_character_,
    observationComments = NA_character_, # manual classification: v_observer_manual and v_type_manual_classification
    .keep = "none"
  ) %>% 
  dplyr::arrange(observationID) %>% 
  dplyr::select(
    observationID, deploymentID, mediaID, eventID, eventStart,
    eventEnd, observationLevel, observationType, cameraSetupType,
    scientificName, count, lifeStage, sex, behavior, individualID,
    individualPositionRadius, individualPositionAngle, individualSpeed,
    bboxX, bboxY, bboxWidth, bboxHeight, classifificationMethod, classififiedBy, classifificationTimestamp,
    classifificationProbability, observationTags, observationComments
  )
```

# Write the three tables

```{r}
write.csv2(deployments, here::here("gbif", "processed", "deployments.csv"))
write.csv2(media, here::here("gbif", "processed", "media.csv"))
write.csv2(observations, here::here("gbif", "processed", "observations.csv"))
```



# Create metadata package

```{r}
acoustic <-
  frictionless::create_package() %>%
  add_resource(resource_name = "deployments", metadata = deployments, schema = here::here("camtrap-dp", "deployments-table-schema-acoustic.json")) %>%
  add_resource(resource_name = "media", metadata = media, schema = here::here("camtrap-dp", "media-table-schema-acoustic.json")) %>%
  add_resource(resource_name = "observations", metadata = observations, schema = here::here("camtrap-dp", "observations-table-schema-acoustic.json")) %>%
  append(c(
    name = "camtrap-dp-acoustic-example-metadataset", # Required
    profile = "https://raw.githubusercontent.com/tdwg/camtrap-dp/1.0.1/camtrap-dp-profile.json", # Required (incorrect, hack to use read_camtrapdp())
    id = NULL,
    created = format(Sys.time(), "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"), # Required
    title = "The Sound of Norway: bird detections",
    description = NULL,
    version = "1.0",
    keywords = c("acoustic", "birds", "passive acoustic monitoring"),
    image = NULL,
    homepage = NULL,
    sources = NULL,
    bibliographicCitation = "Bick, I. Avery, Vegar Bakkestuen, Benjamin Cretois, Ben Hillier, John A. Kålås, Marius Pedersen, Kiran Raja, et al. ‘National-Scale Acoustic Monitoring of Avian Biodiversity and Phenology’. Ecology, 24 May 2024. https://doi.org/10.1101/2024.05.21.595242",
    coordinatePrecision = NULL,
    spatial = NULL, # Required, but keep this empty, this comes later
    temporal = NULL, # Required, but keep this empty, this comes later
    taxonomic = NULL, # Required, but keep this empty, this comes later
    relatedIdentifiers = NULL,
    references = NULL
  )) %>%
  frictionless::create_package()

acoustic$contributors <- list( # Required
  list(
    title = "Julia Wiel",
    email = "julia.wiel@nina.no",
    path = "https://orcid.org/0009-0009-7224-3953",
    role = "contributor",
    organization = "Norwegian Institute for Nature Research (NINA)"
  ),
  list(
    title = "Benjamin Cretois",
    email = "benjamin.cretois@nina.no",
    path = "https://orcid.org/0000-0001-8668-3321",
    role = "rightsHolder",
    organization = "Norwegian Institute for Nature Research (NINA)"
  ),
  list(
    title = "Carolyn Rosten",
    email = "carolyn.rosten@nina.no",
    path = "https://orcid.org/0000-0002-1117-444X",
    role = "rightsHolder",
    organization = "Norwegian Institute for Nature Research (NINA)"
  ),
  list(
    title = "Sarab Sethi",
    email = "sarab.sethi@imperial.ac.uk",
    path = "https://orcid.org/0000-0002-5939-0432",
    role = "principalInvestigator",
    organization = "Imperial College London"
  ),
  list(
    title = "I. Avery Bick",
    email = "i.avery.bick@gmail.com",
    path = "https://orcid.org/0000-0002-8007-9649",
    role = "principalInvestigator",
    organization = "Norwegian Institute for Nature Research (NINA)"
  )
)

acoustic$licenses <- list(
  list(
    name = "CC0-1.0",
    scope = "metadata" # Required
  ),
  list(
    path = "http://creativecommons.org/licenses/by/4.0/",
    scope = "media" # Required
  )
)

acoustic$project <- list( # Required
  # id = "",
  title = "The Sound of Norway", # Required
  # acronym = "",
  description = "The Sound of Norway project monitors ecosystems across the nation with real- time recording cameras. Using machine learning to automatically classifify bird vocalisations, we can simultaneously track the distribution and migration patterns of many species on vast scales and over long time-periods.",
  protocolType = list(c("acoustic")), # Required
  samplingDesign = "systematicRandom", # Required
  path = "https://thesoundofnorway.com/",
  captureMethod = list(c("continuous")), # Required | Bug, see https://github.com/frictionlessmetadata/frictionless-r/issues/276
  individualAnimals = FALSE, # Required
  observationLevel = list(c("media")) # Required | Bug, see https://github.com/frictionlessmetadata/frictionless-r/issues/276
)
```

## Create taxonomy

```{r}
map_taxonomy <- function(taxonomy) {
  list(
    scientificName = taxonomy$scientificName,
    vernacularNames = list(
      eng = taxonomy$EnglishName
      )
  )
}
```

```{r}
taxonomy_list <- apply(taxonomy, 1, as.list)
acoustic$taxonomic <- purrr::map(taxonomy_list, map_taxonomy)
```

# Write metadata Package

```{r}
frictionless::write_package(acoustic, here::here("metadata", "interim", "sound_of_norway"))
```

# Read with camtrapdp

This automatically updates the spatial, temporal and taxonomic coverage in the metametadata, based on the metadata.

```{r}
acoustic_dp <- camtrapdp::read_camtrapdp(here::here("metadata", "interim", "sound_of_norway", "metadatapackage.json"))
```

# Write Camtrap DP

```{r}
camtrapdp::write_camtrapdp(acoustic_dp, here::here("metadata", "processed", "sound_of_norway"))
```
