---
title: "coat_to_camtrapdp"
author: "Julia Wiel"
date: "`r Sys.Date()`"
output: "html_document"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(here)
library(dplyr)
library(tidyr)
library(frictionless)
library(camtrapdp)
library(lubridate)
library(readr)
library(fs)
```

# Import metadata and classifications from borgefjell

```{r}
# Enter variables
year <- 2024

# Import files
setwd("/run/user/210015540/gvfs/sftp:host=login.nird.sigma2.no,user=jwiel/nird/projects/NS11082K/national_monitoring/gbif/")
metadata_raw <- readr::read_delim(here::here("raw", "image_metadata", "national_monitoring_rodents_cameratraps_image_metadata_borgefjell_2024.txt"),
 delim = ";")

classif_raw <- readr::read_delim(here::here("raw", "image_classification_small_mammal_model_2025", "national_monitoring_rodents_cameratraps_image_classification_borgefjell_2024.txt"),
 delim = ";")

coord_raw <- readr::read_delim(here::here("raw", "Koordinates_SN_FI_RE_JO_KO_v2.csv"),
 delim = ";")

```

# Create resources

## Inspect raw data

```{r}
(unique_values_per_column <-
  metadata_raw %>%
  dplyr::summarise(dplyr::across(dplyr::everything(), ~ dplyr::n_distinct(.))) %>%
  tidyr::pivot_longer(dplyr::everything(), names_to = "column", values_to = "n_unique"))

(unique_values_per_column <-
  classif_raw %>%
  dplyr::summarise(dplyr::across(dplyr::everything(), ~ dplyr::n_distinct(.))) %>%
  tidyr::pivot_longer(dplyr::everything(), names_to = "column", values_to = "n_unique"))

(unique_values_per_column <-
  coord_raw %>%
  dplyr::summarise(dplyr::across(dplyr::everything(), ~ dplyr::n_distinct(.))) %>%
  tidyr::pivot_longer(dplyr::everything(), names_to = "column", values_to = "n_unique"))
```

## Format information from metadata

```{r}
metadata <- metadata_raw

# Format captureMethod (per-row)
# Compute captureMethod for each row and store as a character vector
metadata <- metadata %>%
  dplyr::mutate(captureMethod = dplyr::case_when(
    v_sequence == 0 ~ "timeLapse",
    v_sequence > 0  ~ "activityDetection",
    TRUE            ~ NA_character_
  ))

# Format timestamp to UTC
metadata$timestamp <- format(lubridate::ymd_hms(paste(metadata$t_date,metadata$t_time," "), tz = "CEST"), 
format = "%Y-%m-%dT%H:%M:%OSZ")

 # Format deployment dates
 metadata <- metadata %>%
 dplyr::mutate(deploymentStart = min(timestamp),
 deploymentEnd = max(timestamp))

 ## Add coordinates to metadata
 # Transform the key column
coord <- coord_raw
coord <- coord %>%
  dplyr::mutate(sn_site = tolower(name)) %>% 
  mutate(sn_site = gsub("-", "_", sn_site))

# Merge columns from Koordinates into metadata
metadata <- metadata %>% 
  dplyr::left_join(coord, by = "sn_site")

# Create media filePath
metadata <- metadata %>%
  dplyr::mutate(filePath = fs::path("national_monitoring", "images_renamed", sn_locality, sn_site, year,v_image_name))
  #TEST
  #setwd("/run/user/210015540/gvfs/sftp:host=login.nird.sigma2.no,user=jwiel/nird/projects/NS11082K/")
   # file.exists(path_wd(metadata$filePath[5]))

# Create eventID
metadata <- metadata %>%
  dplyr::mutate(eventID = paste(v_trigger_mode, v_event, sep = "_"))

# Create event start and end
metadata <- metadata %>%
  dplyr::group_by(eventID) %>%
  dplyr::mutate(eventStart = min(timestamp),
  eventEnd = max(timestamp)) %>% dplyr::ungroup()
  # Weird: 3310 unique eventID and unique eventStart and only 3305 unique eventEnd

```

## Format information from classifification

```{r}
classif <- classif_raw

# Add observationType
animal <- c("aves", "cricetidae", "lem_lem", "mus_erm", "mus_niv", "sor_sp", "unidentified_animal")
classif <- classif %>%
  dplyr::mutate(observationType = dplyr::case_when(
    v_class_id %in% animal      ~ "animal",
    v_class_id == "bad_quality" ~ "unknown",
    v_class_id == "empty"       ~ "blank",
    TRUE                        ~ "unclassified"
  ))

# Add scientificName
classif <- classif %>%
  dplyr::mutate(scientificName = dplyr::case_when(
    v_class_id == "aves"     ~ "Aves",
    v_class_id == "cricetidae" ~ "Cricetidae",
    v_class_id == "lem_lem"       ~ "Lemmus lemmus",
    v_class_id == "mus_erm"       ~ "Mustela erminea",
    v_class_id == "mus_niv"      ~ "Mustela nivalis",
    v_class_id == "sor_sp"      ~ "Sorex spp.",
    TRUE                        ~ NA_character_
  ))

  # Create observationID
classif <- classif %>%
  dplyr::group_by(v_image_name) %>%
  dplyr::mutate(
    observationID = paste0(v_image_name, "_", row_number())
  ) %>%
  ungroup()

# Add event info from metadata
classif <- classif %>%
  left_join(
    metadata %>% select(v_image_name, eventID, eventStart, eventEnd),
    by = "v_image_name"
  )

# Priorise manual observation and format with automatic
#classif <- classif %>%
 # dplyr::mutate(classificationMethod = )
```

## Create deployments resource_start changes

```{r}
deployments <-
  metadata %>% 
  dplyr::mutate(
    deploymentID = metadata$sn_site, # Required
    locationID = metadata$sc_section,
    locationName = metadata$sn_locality,
    latitude = metadata$lat, # Required
    longitude = metadata$lon, # Required
    coordinateUncertainty = NA_integer_,
    deploymentStart = metadata$deploymentStart, # Required
    deploymentEnd = metadata$deploymentEnd, # Required
    setupBy = NA_character_,
    cameraID = NA_character_,
    cameraModel = NA_character_,
    cameraDelay = 60,
    cameraHeight = NA_real_,
    cameraDepth = NA_real_,
    cameraTilt = -90,
    cameraHeading = NA_integer_,
    detectionDistance = NA_real_,
    timestampIssues = NA,
    baitUse = FALSE,
    featureType = NA_character_,
    habitat = metadata$sc_type_of_sites_ecological,
    deploymentGroups = NA_character_,
    deploymentTags = NA_character_,
    deploymentComments = NA_character_,
    .keep = "none") %>% 
  dplyr::arrange(deploymentID) %>% 
  dplyr::distinct(deploymentID, .keep_all = TRUE) %>%  # Keep only first occurrence of each sn_site
  dplyr::select(
    deploymentID, locationID, locationName, latitude, longitude, 
    coordinateUncertainty, deploymentStart, deploymentEnd, setupBy, 
    cameraID, cameraModel, cameraDelay, cameraHeight, 
    cameraDepth, cameraTilt, cameraHeading, detectionDistance,
    timestampIssues, baitUse, featureType, habitat, deploymentGroups,
    deploymentTags, deploymentComments
  )
```

## Create media resource

```{r}
media <-
  metadata %>% 
  dplyr::mutate(
    mediaID = metadata$v_image_name, # Required
    deploymentID = metadata$sn_site, # Required
    captureMethod = as.character(metadata$captureMethod),
    timestamp = metadata$timestamp, # Required 
    filePath = metadata$filePath,# Required
    filePublic = FALSE, # Required
    fileName = metadata$v_image_name_original,
    fileMediatype = "image/jpeg", # Required
    exifData = NA,
    favorite = NA,
    mediaComments = metadata$v_comment,
    .keep = "none"
  ) %>% 
  dplyr::arrange(mediaID) %>%
  dplyr::distinct() %>% 
  dplyr::select(
    mediaID, deploymentID, captureMethod, timestamp, filePath, 
    filePublic, fileName, fileMediatype, exifData, favorite, mediaComments
  )
```

## Create observations resource

```{r}
observations <-
  classif %>% 
  dplyr::mutate(
    observationID = classif$observationID, # Required
    deploymentID = classif$sn_site, # Required
    mediaID = classif$v_image_name,
    eventID = classif$eventID,
    eventStart = classif$eventStart, # Required
    eventEnd = classif$eventEnd, # Required
    observationLevel = "media", # Required
    observationType = classif$observationType, # Required
    cameraSetupType = NA_character_,
    scientificName = classif$scientificName,
    count = NA_integer_,
    lifeStage = NA_character_,
    sex = NA_character_,
    behavior = NA_character_,
    individualID = NA_character_,
    individualPositionRadius = NA_real_,
    individualPositionAngle = NA_real_,
    individualSpeed = NA_real_,
    bboxX = NA_real_,
    bboxY = NA_real_,
    bboxWidth = NA_real_,
    bboxHeight = NA_real_,
    classificationMethod = "machine", # or split manual and automatic
    classifiedBy = "Small_mammal_classification_model_version_2025",
    classificationTimestamp = NA_Date_,
    classificationProbability = classif$v_confidence_automatic,
    observationTags = NA_character_,
    observationComments = NA_character_, # manual classification: v_observer_manual and v_type_manual_classification
    .keep = "none"
  ) %>% 
  dplyr::arrange(observationID) %>% 
  dplyr::select(
    observationID, deploymentID, mediaID, eventID, eventStart,
    eventEnd, observationLevel, observationType, cameraSetupType,
    scientificName, count, lifeStage, sex, behavior, individualID,
    individualPositionRadius, individualPositionAngle, individualSpeed,
    bboxX, bboxY, bboxWidth, bboxHeight, classificationMethod,
    classifiedBy, classificationTimestamp, classificationProbability,
    observationTags, observationComments
  )
```

# Write the three tables

```{r}
write.csv2(deployments, here::here("processed", "deployments.csv"))
write.csv2(media, here::here("processed", "media.csv"))
write.csv2(observations, here::here("processed", "observations.csv"))
```

## Check that the three tables conform to the Camtrap DP schemas
```{r}
# --- Compare data frames to Camtrap DP schemas ---
library(jsonlite)

# Compare deployments
deployments_schema_url <- "https://raw.githubusercontent.com/tdwg/camtrap-dp/1.0/deployments-table-schema.json"
deployments_schema <- fromJSON(deployments_schema_url)
deployments_expected <- deployments_schema$fields$name
deployments_actual <- colnames(deployments)
cat("\nDeployments table:\n")
cat("Missing fields: ", paste(setdiff(deployments_expected, deployments_actual), collapse = ", "), "\n")
cat("Extra fields: ", paste(setdiff(deployments_actual, deployments_expected), collapse = ", "), "\n")

# Compare media
media_schema_url <- "https://raw.githubusercontent.com/tdwg/camtrap-dp/1.0/media-table-schema.json"
media_schema <- fromJSON(media_schema_url)
media_expected <- media_schema$fields$name
media_actual <- colnames(media)
cat("\nMedia table:\n")
cat("Missing fields: ", paste(setdiff(media_expected, media_actual), collapse = ", "), "\n")
cat("Extra fields: ", paste(setdiff(media_actual, media_expected), collapse = ", "), "\n")

# Compare observations
observations_schema_url <- "https://raw.githubusercontent.com/tdwg/camtrap-dp/1.0/observations-table-schema.json"
observations_schema <- fromJSON(observations_schema_url)
observations_expected <- observations_schema$fields$name
observations_actual <- colnames(observations)
cat("\nObservations table:\n")
cat("Missing fields: ", paste(setdiff(observations_expected, observations_actual), collapse = ", "), "\n")
cat("Extra fields: ", paste(setdiff(observations_actual, observations_expected), collapse = ", "), "\n")
```

# Create metadata package

```{r}
properties <-
  frictionless::create_package() %>%
  add_resource(resource_name = "deployments", data = deployments, schema = "https://raw.githubusercontent.com/tdwg/camtrap-dp/1.0/deployments-table-schema.json") %>%
  add_resource(resource_name = "media", data = media, schema = "https://raw.githubusercontent.com/tdwg/camtrap-dp/1.0/media-table-schema.json") %>%
  add_resource(resource_name = "observations", data = observations, schema = "https://raw.githubusercontent.com/tdwg/camtrap-dp/1.0/observations-table-schema.json") %>%
  append(c(
    profile = "https://raw.githubusercontent.com/tdwg/camtrap-dp/1.0/camtrap-dp-profile.json", # Required modified from 1.0.1 to 1.0 but no effect in .json
    name = "camtrap-dp-rodent-example-metadataset", # Required
    id = NULL,
    created = format(Sys.time(), "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"), # Required
    title = "National monitoring of rodents with camera traps in Borgefjell, Norway", # Required
    description = NULL,
    version = "1.0",
    keywords = c("rodent", "camera trap", "national monitoring"),
    image = NULL,
    homepage = NULL,
    sources = NULL,
    bibliographicCitation = NULL,
    coordinatePrecision = NULL,
    spatial = NULL, # Required, but keep this empty, this comes later
    temporal = NULL, # Required, but keep this empty, this comes later
    taxonomic = NULL, # Required, but keep this empty, this comes later
    relatedIdentifiers = NULL,
    references = NULL
  )) %>%
  frictionless::create_package()
```

```{r}

properties$contributors <- list( #Required
  list(
    title = "Eivind Flittie Kleiven",
    email = "eivind.kleiven@nina.no",
    path = "https://orcid.org/0000-0002-9556-0195",
    role = "principalInvestigator",
    organization = "Norwegian Institute for Nature Research (NINA)"
  )
)

properties$project <- list( #Required
  title = "National rodent monitoring progam, Norway",
  samplingDesign = "systematicRandom",
  captureMethod = c("activityDetection", "timeLapse"),
  individualAnimals = FALSE,
  observationLevel = c("media", "event")
)
```

# Write metadata Package

```{r}
frictionless::write_package(properties, here::here("interim"))
```

# Read with camtrapdp

This automatically updates the spatial, temporal and taxonomic coverage in the metametadata, based on the metadata.

```{r}
properties_dp <- camtrapdp::read_camtrapdp(here::here("interim", "datapackage.json"))
```

# Write Camtrap DP

```{r}
camtrapdp::write_camtrapdp(properties_dp, here::here("processed"))
```

# Fix Camtrap DP JSON errors
```{r}
# Fix the schema references that got corrupted by camtrapdp processing
final_package <- jsonlite::fromJSON(here::here("processed", "datapackage.json"))

# Replace the schema arrays with simple URL strings
# The resources is a data.frame, so we access the schema column directly
final_package$resources$schema[[1]] <- "https://raw.githubusercontent.com/tdwg/camtrap-dp/1.0/deployments-table-schema.json"
final_package$resources$schema[[2]] <- "https://raw.githubusercontent.com/tdwg/camtrap-dp/1.0/media-table-schema.json"
final_package$resources$schema[[3]] <- "https://raw.githubusercontent.com/tdwg/camtrap-dp/1.0/observations-table-schema.json"

# Fix the profile version that gets changed from 1.0 to 1.0.1
final_package$profile <- "https://raw.githubusercontent.com/tdwg/camtrap-dp/1.0/camtrap-dp-profile.json"

# Write the corrected package back to file
jsonlite::write_json(final_package, here::here("processed", "datapackage.json"), pretty = TRUE, auto_unbox = TRUE)
```
